---
title: "Sampling and Summarizing Data"
author: "Eric R. Scott, Avalon Owens"
output:
  ioslides_presentation:
  pdf_document: default
---

## Overview
- Summarizing data by calculating summary statistics with the `dplyr` package
    + `summarize()`, `group_by()`, `mean()`, `median()`, `var()`, `sd()`, `min()`, `max()`, `round()`
- More plots with `ggplot2` (box plots, violin plots, jitter plots, combining plots)
- Sampling with R (`c()`, `rep()`, and `sample()`)


# Coding tip #1

## Named and unnamed arguments
- So far, we've been using explicitly named arguments for all the functions
    + E.g: `ggplot(`**`data = iris,`**` aes(x = Sepal.Length, y = Sepal.Width))`
- You don't have to name them (the `data = ` part) if you supply them in the right order.
    + `ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width))` works just fine.
    + To figure out the right order, check the example in the help file.
    

## Summarizing data
- We're going to use the `dplyr` package for summarizing data

`summarise(<<data frame>>, <<new variable>> = <<summary function>>)`

```{r message=FALSE, warning=FALSE}
library(dplyr)
summarize(iris, mean_sep.len = mean(Sepal.Length))
```

## Making a summary table

```{r}
summarise(iris,
          mean = mean(Sepal.Length),
          median = median(Sepal.Length),
          range = max(Sepal.Length) - min(Sepal.Length),
          variance = var(Sepal.Length))
```

## Summarizing within groups
- Sometimes it's useful to get summary statistics for groups separately, just like the faceted histograms we made before.
- use `group_by()` to add *invisible* grouping variable

```{r}
iris_grouped <- group_by(iris, Species)

summarise(iris_grouped,
          mean = mean(Sepal.Length),
          median = median(Sepal.Length),
          range = max(Sepal.Length) - min(Sepal.Length),
          variance = var(Sepal.Length))
```

# Coding tip #2

## The "pipe" ( %>% )

- Used in `dplyr` and many other Tidyverse packages
- Keyboard shortcut: command/control (macOS/Windows) + shift + m
- Allows you to "chain" functions together

```{r}
iris %>%
  group_by(Species) %>%
  summarize(mean = mean(Sepal.Length))
```

## Readable code

- Using the pipe makes your code more readable (to humans!)

```{r}
iris %>%
  group_by(Species) %>%
  summarize(mean = mean(Sepal.Length))
```

"Take the iris dataset, group it by Species, then summarize it by taking the mean of Sepal Length for each Species"

## Under the hood of `%>%`

- What it actually does is put whatever is on it's left into the first argument of the function on it's right
```{r eval = FALSE}
iris %>% group_by(Species)
# is the same as
group_by(iris, Species)
```
- It only works with functions that have data as their first argument (all Tidy verse functions, not all base R functions)


## Practice
- Make this code more readable using ` %>% `
```{r eval=FALSE}
iris_grouped <- group_by(iris, Species)

summarize(iris_grouped,
          mean = mean(Sepal.Length),
          median = median(Sepal.Length),
          range = max(Sepal.Length) - min(Sepal.Length),
          variance = var(Sepal.Length))
```


